Certified Knowledge Compilation \\ with Application to Verified Model Counting

### 2023-03-15 doc/mc-paper.tex
orcid
RUP

Performing ring evaluation on an arbitrary Boolean formula could be intractable, but it is straightforward for a formula with partitioned operations:
\begin{proposition}
Ring evaluation with operations $\boolnot$, $\pand$, and $\por$ satisfies the following for any labeling $w$:
\begin{eqnarray}
\rep(\boolnot \phi, w) &=& \mulident - \rep(\phi, w) \label{eqn:ring:negation} \\
\rep\left(\Pand_{i=1,k} \phi_i, w \right) &=& \prod_{i=1,k} \rep(\phi_i, w) \label{eqn:ring:product} \\
\rep\left(\Por_{i=1,k} \phi_i, w\right) &=& \sum_{i=1,k} \rep(\phi_i, w) \label{eqn:ring:sum}
\end{eqnarray}
\end{proposition}

@@

>>>>

Performing ring evaluation on an arbitrary Boolean formula could be intractable, but it is straightforward for a formula with partitioned operations:
\begin{proposition}
Ring evaluation with operations $\boolnot$, $\pand$, and $\por$ satisfies the following for any labeling $w$:
\begin{eqnarray}
\rep(\boolnot \phi, w) &=& \mulident - \rep(\phi, w) \label{eqn:ring:negation} \\
\rep\left(\phi_1, w \pand \dots \pand \phi_k, w \right) &=& \prod_{i=1..k} \rep(\phi_i, w) \label{eqn:ring:product} \\
\rep\left(\phi_1, w \por \dots \por \phi_k, w \right) &=& \sum_{i=1..k} \rep(\phi_i, w) \label{eqn:ring:sum}
%\rep\left(\Pand_{i=1,k} \phi_i, w \right) &=& \prod_{i=1,k} \rep(\phi_i, w) \label{eqn:ring:product} \\
%\rep\left(\Por_{i=1,k} \phi_i, w\right) &=& \sum_{i=1,k} \rep(\phi_i, w) \label{eqn:ring:sum}
\end{eqnarray}
\end{proposition}

@@

As an example, consider the formula $\varphi = (\obar{a} \lor b) \land (\obar{b} \lor c \lor \obar{e}) \land (\obar c \lor d)$ 
and a RUP step to derive the target clause $(\obar{a} \lor d \lor \obar{e})$ from the three clauses from $\varphi$.
A RUP proof would take the following form.  We start with the assignment that falsifies the target clause. This assignment
is extended by clauses in $\varphi$ that become unit and eventually the empty clause. In the illustration below, the 
involved clauses (antecedents) are shown in the order that they became unit / falsified.

\begin{center}
  \begin{tabular}{lcccc}
         & \makebox[15mm]{Target}    & \makebox[10mm]{} & \makebox[10mm]{Antecedents} & \makebox[10mm]{} \\
  Clause & $\obar{a} \lor d \lor \obar{e}$ & $\obar{c} \lor d$  & $\obar{e} \lor \obar{b} \lor c$   & $\obar{a} \lor b$ \\
  \midrule
  Units  & $a$, $\obar{d}$, $e$      &  $\obar{c}$             & $\obar{b}$                         & $\nil$ \\
  \end{tabular}
\end{center}

RUP is an alternative formulation of resolution.  For target clause
$C$, it can be seen that applying resolution operations to the
antecedent clauses from right to left will derive a clause $C'$ such
that $C' \subseteq C$.  By {\em subsumption} %~\cite{philipp:lpar:2017},
we then have $C' \rightarrow C$.  Compared to listing each resolution
operation as a separate step, using RUP as the basic proof step makes
the proofs more compact.
